Geometrization of characters of tori over non-Archimedean local fields

### 2013-07-19 shanghai.tex
Changing some margin notes

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' in this
paper. Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$ \David{I think our notation is clearer if we use $k$
  rather than $\Fq$}; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,F)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob_{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob_{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

@@

The main result of this paper is a canonical isomorphism
\[
\CSiso{\mathswab{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of character sheaves on
$\mathswab{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \Clifton{I'll come back to working on
  this when we're closer to being finished. I'd be just as happy if
  you took this cup from my lips, though.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasi-characters of
%$K^\times$: the monoidal category of character sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

@@

  The fact that the functor $f^* : \CS(G) \to \CS(H)$ commutes with
  the trace of Frobenius, in the sense above, follows immediately from
  the definitions; see \cite{Laumon}*{1.1.1.2}, where the ambient
  hypothesis that $X$ is of finite type over $\Fq$ can be replaced by
  the hypothesis that $X$ is locally of finite type over $\Fq$.
  \Clifton{I have not proved $(f\circ g)^* = g^* \circ f^*$, yet.}
\end{proof}

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi-2}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
\[
\begin{tikzcd}
\CS(0) \rar & \CS(\pi_0(G)) \rar & \CS(G) \rar & \CS(G^0) \rar & \CS(0)
\end{tikzcd}
\]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
\begin{equation}\label{eq:pi-1}
\begin{tikzcd}
\CSiso{0} \rar & \CSiso{\pi_0(G)} \rar & \CSiso{G} \rar & \CSiso{G^0} \rar & \CSiso{0}.
\end{tikzcd}
\end{equation}
It is easy to see that $\CSiso{0}$ is trivial. In this paper we will find $\CSiso{G^0}$ and $\CSiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated, and we will also see that \eqref{eq:pi-1} is split exact; this allows us to find $\CSiso{G}$. To prove this, we find it useful to offer an alternate perspective on a subcategory of $\CS(G)$, in the next section.

@@

  To see why this is so, consider the category \cdef{$\CS_0(G)$} of
  pairs $(\cs{L}_0,\mu_0)$ where $\cs{L}_0$ is a local system on $G$
  and $\mu_0: m^*\cs{L}_0 \to \cs{L}_0 \tight{0}{\boxtimes}{0} \cs{L}_0$ satisfies
  condition \ref{CS.1} adapted to sheaves on $G$; maps in $\CS_0(G)$
  are defined by the second part of condition \ref{CS.2}, {\it mutatis mutandis}.
  \David{Should we move this definition since it will be important elsewhere?}
  \Clifton{Agreed. I'll do that.}
  Let $b : \bG \to G$ be base change
  (pull-back of $G\to \Spec{\Fq}$ along $\Spec{\bFq} \to \Spec{\Fq}$).
  The proof of Proposition~\ref{prop:pullback} adapts
  (simplifies, in fact) to show that $b^* : \CS_0(G) \to \CS(G)$ is a
  monoidal functor. This functor is fully faithful; see
  \cite{SGA7.2}*{Expos\'e XIII} and \cite{BBD}*{Prop. 5.2.1}.  We will
  show that $b^*$ is essentially surjective.

@@

Let $G^0$ be the connected component of $G$ containing the identity.
By Proposition~\ref{prop:pullback}, pull-back along
$G^0\hookrightarrow G$ determines a monoidal functor
$\cs{L}\mapsto \cs{L}\vert_{G^0}$ from $\CS(G)$ to $\CS(G^0)$.  In
this section we show that this functor is essentially surjective. To
do this we employ a perspective on characters sheaves which is
different from that in Section~\ref{section:category} but which works
very well for geometrically connected abelian algebraic groups; this
perspective fails to extend properly to the more general context of
Section~\ref{section:category}.

>>>>

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' in this
paper. Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,F)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob_{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob_{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

@@

The main result of this paper is a canonical isomorphism
\[
\CSiso{\mathswab{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of character sheaves on
$\mathswab{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \David{I'll do this when we're closer to being finished.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasi-characters of
%$K^\times$: the monoidal category of character sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

@@

  The fact that the functor $f^* : \CS(G) \to \CS(H)$ commutes with
  the trace of Frobenius, in the sense above, follows immediately from
  the definitions; see \cite{Laumon}*{1.1.1.2}, where the ambient
  hypothesis that $X$ is of finite type over $\Fq$ can be replaced by
  the hypothesis that $X$ is locally of finite type over $\Fq$.

  Finally, the fact that $(f\circ g)^* = g^* \circ f^*$ follows from the analogous
  statements about pullbacks of sheaves and maps of sheaves.
  \David{Is this explanation of $(f\circ g)^* = g^* \circ f^*$ sufficient?}
\end{proof}

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi-2}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
\[
\begin{tikzcd}
\CS(0) \rar & \CS(\pi_0(G)) \rar & \CS(G) \rar & \CS(G^0) \rar & \CS(0)
\end{tikzcd}
\]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
\begin{equation}\label{eq:pi-1}
\begin{tikzcd}
\CSiso{0} \rar & \CSiso{\pi_0(G)} \rar & \CSiso{G} \rar & \CSiso{G^0} \rar & \CSiso{0}.
\end{tikzcd}
\end{equation}
It is easy to see that $\CSiso{0}$ is trivial. In this paper we will find $\CSiso{G^0}$ and $\CSiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated, and we will also see that \eqref{eq:pi-1} is split exact \David{is it really split?}; this allows us to find $\CSiso{G}$. To prove this, we find it useful to offer an alternate perspective on a subcategory of $\CS(G)$, in the next section.

@@

  To see why this is so, consider the category \cdef{$\CS_0(G)$} of
  pairs $(\cs{L}_0,\mu_0)$ where $\cs{L}_0$ is a local system on $G$
  and $\mu_0: m^*\cs{L}_0 \to \cs{L}_0 \tight{0}{\boxtimes}{0} \cs{L}_0$ satisfies
  condition \ref{CS.1} adapted to sheaves on $G$; maps in $\CS_0(G)$
  are defined by the second part of condition \ref{CS.2}, {\it mutatis mutandis}.
  \David{This section should be modified once we're done with the section on bounded character sheaves}
  Let $b : \bG \to G$ be base change
  (pull-back of $G\to \Spec{\Fq}$ along $\Spec{\bFq} \to \Spec{\Fq}$).
  The proof of Proposition~\ref{prop:pullback} adapts
  (simplifies, in fact) to show that $b^* : \CS_0(G) \to \CS(G)$ is a
  monoidal functor. This functor is fully faithful; see
  \cite{SGA7.2}*{Expos\'e XIII} and \cite{BBD}*{Prop. 5.2.1}.  We will
  show that $b^*$ is essentially surjective.

@@

\David{This section needs to be modified once the bounded character sheaves section is done}
Let $G^0$ be the connected component of $G$ containing the identity.
By Proposition~\ref{prop:pullback}, pull-back along
$G^0\hookrightarrow G$ determines a monoidal functor
$\cs{L}\mapsto \cs{L}\vert_{G^0}$ from $\CS(G)$ to $\CS(G^0)$.  In
this section we show that this functor is essentially surjective. To
do this we employ a perspective on characters sheaves which is
different from that in Section~\ref{section:category} but which works
very well for geometrically connected abelian algebraic groups; this
perspective fails to extend properly to the more general context of
Section~\ref{section:category}.
